Problem Set 3: Part 1
========================================================
- Chad Dunham

Version Check:
print(version$version.string)

Packages Used

randomForest: 4.6.7

party: 1.0.13



Installing needed packages

```{r}
install.packages("randomForest", dependencies=T)
library(randomForest)

install.packages("party", dependencies=T)
library(party)

#Version Check
packageVersion("randomForest")
packageVersion("party")
```

Import the data

```{r}
titanicData = read.csv("titanic3.csv")
head(titanicData)
```

Creating the training and testing data

```{r}

sampleSplit = sample(2, nrow(titanicData), replace=T, prob=c(.7, .5))
titanicTrain = titanicData[sampleSplit==1,]
titanicTest = titanicData[sampleSplit==2,]

```

======= TRAINING DATA SECTION =======

Run glm of Titanic Data

```{r}
titanicSurvivalTrain = glm(survived ~ pclass + sex + pclass:sex + age + sibsp, family = binomial(logit), data=titanicTrain)

summary(titanicSurvivalTrain)
```


Random Forest of Titanic Data

```{r}
titanicRandomForestTrain = randomForest(as.factor(survived) ~ pclass + sex + age + sibsp, data=titanicTrain, ntree=5000, importance = T, na.action = na.omit)

titanicRandomForestTrain

plot(titanicRandomForestTrain)
```

Conditional Tree Model of Titanic Data

```{r}
titanicCondTreeTrain = ctree(as.factor(survived) ~ pclass + sex + age + sibsp, data=titanicTrain)

titanicCondTreeTrain

plot(titanicCondTreeTrain)
```


======= TEST DATA SECTION =======

Run glm of Titanic Data

```{r}
titanicSurvivalTest = glm(survived ~ pclass + sex + pclass:sex + age + sibsp, family = binomial(logit), data=titanicTest)

summary(titanicSurvivalTest)
```


Random Forest of Titanic Data

```{r}
titanicRandomForestTest = randomForest(as.factor(survived) ~ pclass + sex + age + sibsp, data=titanicTest, ntree=5000, importance = T, na.action = na.omit)

titanicRandomForestTest
plot(titanicRandomForestTest)

```

Conditional Tree Model of Titanic Data

```{r}
titanicCondTreeTest = ctree(as.factor(survived) ~ pclass + sex + age + sibsp, data=titanicTest)

titanicCondTreeTest
plot(titanicCondTreeTest)

```

Confusion Matrices

=== Training Data ===
## Confusion matrix:
##     0   1 class.error
## 0 331  53      0.1380
## 1  84 159      0.3457

=== Test Data ===
## Confusion matrix:
##     0   1 class.error
## 0 214  21     0.08936
## 1  66 118     0.35870

