Project 3, Part 2
- Chad Dunham
==================================================


Packages and Versions Used:

R: 3.0.2

plyr: 1.8.1

lattice: 0.20.27

xtable: 1.7.3



Version Check:
```{r}
print(version$version.string)
```


Loading and verifying the Gapminder data

```{r}
gdURL = "http://www.stat.ubc.ca/~jenny/notOcto/STAT545A/examples/gapminder/data/gapminderDataFiveYear.txt"
gDat = read.delim(file = gdURL)

str(gDat)
```

Data aggregation using ddply

```{r}
(snippet = subset(gDat, country == "Canada"))

install.packages("plyr", dependencies = T)
library(plyr)

#Version Check
packageVersion("plyr")

(maxLeByCont = ddply(gDat, ~ continent, summarize, maxLifeExp = max(lifeExp)))

str(maxLeByCont)
levels(maxLeByCont$continent)
```

GDP per capita function (code not included on website)

```{r}
(minGDPPerCap = ddply(gDat, ~ continent, summarize, minGDP = min(gdpPercap)))

```

Continuing using plyr

```{r}
ddply (gDat, ~continent, summarize, nUniqCountries = length(unique(country)))

ddply(gDat, ~ continent, function(x) return(c(nUniqCountries = length(unique(x$country)))))

ddply(gDat, ~ continent, summarize, minLifeExp = min(lifeExp), maxLifeExp = max(lifeExp), medGdpPercap = median(gdpPercap))

```

Using ddply to fit a linear regression modeling life expectancy

```{r}

install.packages("lattice")
library(lattice)
#Version Check
packageVersion("lattice")


jCountry = "France"
(jDat = subset(gDat, country == jCountry))

xyplot(lifeExp ~ year, jDat, type = c("p", "r"))

jFit = lm(lifeExp ~ year, jDat)
summary(jFit)

(yearMin = min(gDat$year))

jFit = lm(lifeExp ~ I(year - yearMin), jDat)
summary(jFit)

class(jFit)
mode(jFit)

str(jFit)
names(jFit)
jFit$coefficients
coef(jFit)

jFun = function(x) coef(lm(lifeExp ~ I(year - yearMin), x))
jFun(jDat)

jFun = function(x) {
  estCoefs = coef(lm(lifeExp ~ I(year - yearMin), x))
  names(estCoefs) = c("intercept", "slope")
  return(estCoefs)
}
jFun(jDat)
```

Testing the new function

```{r}
jFun(subset(gDat, country == "Canada"))

jFun(subset(gDat, country == "Uruguay"))

jFun(subset(gDat, country == "India"))

```

Finalizing the function

```{r}
jCoefs = ddply(gDat, ~country, jFun)
str(jCoefs)

tail(jCoefs)

```

Reproducing the function in concise form:

```{r}
gdURL = "http://www.stat.ubc.ca/~jenny/notOcto/STAT545A/examples/gapminder/data/gapminderDataFiveYear.txt"
gDat = read.delim(file = gdURL)
yearMin = min(gDat$year)
jFun = function(x) {
  estCoefs = coef(lm(lifeExp ~ I(year - yearMin), x))
  names(estCoefs) = c("intercept", "slope")
  return(estCoefs)
}
jCoefs = ddply(gDat, ~country, jFun)
```

Using the xtable package make a nice table

```{r results="asis"}

install.packages("xtable", dependencies = T)
library(xtable)

#Version Check
packageVersion("xtable")

set.seed(916)
foo = jCoefs[sample(nrow(jCoefs), size = 15), ]
foo = xtable(foo)
print(foo, type = "html", include.rownames = F)

```

Enhancing the ddply call and making a better table

```{r}
jCoefs = ddply(gDat, ~country + continent, jFun)
str(jCoefs)

tail(jCoefs)

```

```{r results="asis"} 
set.seed(916)
foo = jCoefs[sample(nrow(jCoefs), size = 15), ]
foo = arrange(foo, intercept)
foo = xtable(foo)
print(foo, type = "html", include.rownames = F)
```

=================================================
Q & A Section
=================================================

```{r}
jFunTwoArgs = function(x, cvShift = 0) {
  estCoefs = coef(lm(lifeExp ~ I(year - cvShift), x))
  names(estCoefs) = c("intercept", "slope")
  return(estCoefs)
}

jCoefsSilly = ddply(gDat, ~country, jFunTwoArgs)
head(jCoefsSilly)

jCoefsSane = ddply(gDat, ~country, jFunTwoArgs, cvShift = 1952)
head(jCoefsSane)

jCoefsBest = ddply(gDat, ~country, jFunTwoArgs, cvShift = min(gDat$year))
head(jCoefsBest)
```
