Problem Set 3, Part 3
========================================================
- Chad Dunham

Packages and Versions Used:

R: 3.0.2

cluster: 1.15.2

Rattle: 3.0.2

NcClust: 1.4

flexclust: 1.3.4


```{r}
print(version$version.string)
```

Heirarchal Clustering Example
=============================


Installing required packages

```{r}

install.packages("cluster", dependencies=T)
library(cluster)

#Version check
packageVersion("cluster")

```

Importing and quickly looking at the information

```{r}

cars = read.delim('cars.tab', stringsAsFactors=F)
head(cars)

```

Trim the data.fram and standardize the data

```{r}

cars.use = cars[,-c(1,2)]
medians = apply(cars.use, 2, median)
mads = apply(cars.use, 2, mad)
cars.use = scale(cars.use, center=medians, scale=mads)

```

Calculating a distance (Euclidean) matrix and creating a dendogram

```{r}

cars.dist = dist(cars.use)
cars.hclust = hclust(cars.dist)
plot(cars.hclust, labels = cars$Car, main = "Default from hclus")

```

Analyzing the cluster

```{r}

groups.3 = cutree(cars.hclust, 3)
table(groups.3)

counts = sapply(2:6, function(ncl) table(cutree(cars.hclust, ncl)))
names(counts) = 2:6
counts

cars$Car[groups.3==1]

sapply(unique(groups.3), function(g) cars$Car[groups.3 == g])

groups.4 = cutree(cars.hclust, 4)
sapply(unique(groups.4), function(g) cars$Car[groups.4 == g])

table(groups.3, cars$Country)

aggregate(cars.use, list(groups.3), median)

aggregate(cars[,-c(1,2)], list(groups.3), median)

a3 = aggregate(cars[,-c(1,2)], list(groups.3), median)
data.frame(Cluster=a3[,1], Freq=as.vector(table(groups.3)), a3[,-1])

a4 = aggregate(cars[,-c(1,2)], list(groups.4), median)
data.frame(Cluster=a4[,1], Freq=as.vector(table(groups.4)), a4[,-1])

```



K-Means Clustering Example
=============================

Installing required packages

```{r}
install.packages("rattle")
library(rattle)

#version check
packageVersion("rattle")


install.packages("NbClust")
library(NbClust)

#version check
packageVersion("NbClust")


install.packages("flexclust")
library(flexclust)

#version check
packageVersion("flexclust")

```


Setting up functions and standardizing data

```{r}

wssplot = function(data, nc=15, seed=1234) {
  wss = (nrow(data)-1)*sum(apply(data,2,var))
  for (i in 2:nc) {
    set.seed(seed)
    wss[i] = sum(kmeans(data, centers=i)$withinss)}
  plot(1:nc, wss, type="b", xlab="Number of Clusters", ylab="Within groups sum of squares")
  }

data(wine, package="rattle")
head(wine)

df = scale(wine[-1])

```

Determing Number of Clusters

```{r}

wssplot(df)
set.seed(1234)
nc = NbClust(df, min.nc=2, max.nc=15, method="kmeans")
table(nc$Best.n[1,])

barplot(table(nc$Best.n[1,]), xlab = "Number of Clusters", ylab = "Number of Criteria", main = "Number of Clusters Chosen by 26 Criteria")

```


K-Means Cluster Analysis

```{r}

set.seed(1234)

fit.km = kmeans(df, 3, nstart=25)
fit.km$size

fit.km$centers

aggregate(wine[-1], by=list(cluster=fit.km$cluster), mean)

```

Checking the effectiveness of the K-means Cluster

```{r}

ct.km = table(wine$Type, fit.km$cluster)
ct.km

randIndex(ct.km)

```
